---
title: Concept Mover’s Distance
author: "Edoardo Costantini"
output:
  html_document:
    toc: true                    # table of content true
    toc_depth: 4                 # number of headings for toc
    number_sections: true        # if you want number sections at each table header
    highlight: tango             # specifies the syntax highlighting style
    css: ../../css/rmdstyles.css # css file for this project
---

# Preamble

- Project:   knowledgeBase
- Objective: TODO
- Author:    Edoardo Costantini
- Created:   2022-03-10
- Modified:  `r format(Sys.time(), '%Y-%m-%d')`

# Content

Let's see how to asses the a text's engagement with a focal concept using [Concept Mover’s Distance (CMD)](https://www.dustinstoltz.com/code).
For this we will use the `text2map` R package, so be sure to install it before we start.
We will also use the `googledrive` package to download some files we need.

```{r Set up, results = 'hide', message = FALSE, warning = FALSE}
# Set up -----------------------------------------------------------------------

# Load packages
library(tm)
library(text2map)
library(googledrive)

```
Next, let's define a body of text we want to work with.
In my research line, I'm interested in how some topics are dicussed in papers published in sociological journals.
For this example, we will try to analyse three open-source articles from a leading sociological journal.
So to start things off, let's read in the text of three articles from the American Sociological Review (ASR):

```{r prep dtm, results = 'hide', message = FALSE, warning = FALSE}
# Load 20 news articles from tm pacakge dealing with the topic of "crude oil"
?crude
data("crude")

# Create the Document Term Matrix for this corpus
dtm <- DocumentTermMatrix(crude)

# These are the articles ids
dtm$dimnames$Docs

# Check out the terms
head(dtm$dimnames$Terms)

```

Our goal is to find which articles actually deal with the topic of "crude oil".
We want to use the Concept Mover’s Distance to find out.
For a CMD to work, we need to have a good pre-trained source of **word embedding vectors** (with the “words” as row names).
These can usually be found online quite easily.
For a list of ideas look [here]{https://culturalcartography.gitlab.io/text2map/articles/CMDist-concept-movers-distance.html#word-embedding-matrix}.
Here, we will use  fastText English Word Vectors trained on the Common Crawl (which you can download [ft.cc.en.300D.2M.Rds (1.4 Gbs)]{https://drive.google.com/file/d/17H4GOGedeGo0urQdDC-4e5qWQMeWLpGG/view?usp=sharing}).
Store the file in a location you can easily access.
I stored it in a "rds" directory located at the same level as the R script I'm suing for this post.
Then, you can load the file with the following line (but replace `./rds/ft.cc.en.300D.2M.Rds` with the path to your file.
The file is huge.
It will take time.
Be patient.
Take a sip.

```{r load word vector}
# Read the word vector RDS file from your computer
wv <- readRDS("./rds/ft.cc.en.300D.2M.Rds")

# It's just a matrix
class(wv)

# with a LOT of rows
dim(wv)

# Check the first few rows and columns
wv[1:10, 1:5]

```

Looks good.
Now, we want to find words that relate to the topic of interest: crude oil.
We want to find good **anchor terms** for our concept.
We can use the `text2vec::sim2()` function to get the cosine distance between “oil” and its top 50 nearest neighbors.

```{r define anchor terms}
# Compute the cosine distance between survey and all other words in wv
cosSim_survey <- text2vec::sim2(x = wv,
                                y = wv["oil", , drop = FALSE],
                                method = "cosine")

# Print the first 50 nearest neighbors to survey
head(sort(cosSim_survey[, 1], decreasing = TRUE), 50)

```

With a quick scan of these 50 terms I decide that my anchor terms should be:

- oil
- gas
- petroleum
- petrolium

Now that we have settled on the anchor terms, we can use `CMDist` to find the closeness of the documents stored in a DTM matrix to the focal concepts denoted by these words.

```{r CDM analysis single word}
# CDM analysis -----------------------------------------------------------------

# Single Word CDM
closeness_oil <- CMDist(dtm = dtm, cw = "oil", wv = wv)

# Print the
head(closeness_oil)

```

In the output, we can see fr each document ("doc_id") what is the CMD engagement corresponding to the word "oil".

```{r CDM analysis anchors}
# Define a vector of anchors
anchors <- c("oil", "gas", "petroleum", "petrolium")

# Compute the CDM for all of them
closeness_anchors <- CMDist(dtm = dtm, cw = anchors, wv = wv)

# Print the first 6 docs's CDMs
head(closeness_anchors)

```

# TL;DR, just give me the code!
```{r TLDR, ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE}
```