---
title: Moderation in linear models
author: "Edoardo Costantini"
output:
  html_document:
    toc: true                    # table of content true
    toc_depth: 4                 # number of headings for toc
    number_sections: true        # if you want number sections at each table header
    highlight: tango             # specifies the syntax highlighting style
    css: ../../css/rmdstyles.css # css file for this project
---

# Preamble

- Project:   knowledgeBase
- Objective: Showcase moderation in linear models
- Author:    Edoardo Costantini
- Created:   2022-03-09
- Modified:  `r format(Sys.time(), '%Y-%m-%d')`

# Content

## Concepts

When analysis a phenomena with **additive models** we are trying to understand if the (partial) effect
of a predicto on an outcome.
We assume that this partial effect does not change based on the values of the other predictors.
Consider the following linear additive model:
\[
  y = \beta_0 + \beta_1 * x_1 + \beta_2 * x_2
\]
If a unit increase in hours of study ($x_1$) leads to a half-unit increase ($\beta_1$) in the grade of the exam ($y$), controlling for the age of a person ($x_2$), we do not expect $\beta_1$ to be different for students of different ages.
We expect $\beta_1$ to be the same for a 20-year-olds and for a 60-year-olds student.

However, it might be that a single additional hour of study is more impactful for a young student than for an older student.
If this is what you expect, then you want to move away from additive models and start working with **moderation**.
Models with moderation effects allow us to entertain the possibility that the effect of a predictor $x_1$ is different for different levels of another predictor $x_2$.
If this is the case, then we start thinking about *conditional effects*.
What is the effect of an increased unit increase in hours of study ($x_1$) on the exam grade ($y$), for a 20-year-olds? What is that effect for a 60-year-olds?

## R Code

First let's set up the R environment.

```{r Set up, results = 'hide', message = FALSE, warning = FALSE}
# Load packages

```


*Here is what to do*

```{r code chunk name}
# Some code
# Always levea a line empty at the end

```

# TL;DR, just give me the code!
```{r TLDR, ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE}
```